On change of priors in one-vs-one classification methods

### 2025-05-08 paper.tex
Restructure paper into multiple files

\newtheorem{prop}{Proposition}
\newtheorem{thm}{Theorem}
\newtheorem{assumption}{Assumption}

@@

\section{Theoretical basis of coupling methods}

Suppose we have $K$ classes of objects, each distributed with density $d_i$, $i=1,\ldots, K$ on a feature space $X$. In a multi-class classification problem we are provided with the prior probabilities $\Pi = (P_1, \ldots, P_k)$ of each class. Given $x$ in $X$ we aim to find the probability that $x$ belongs to $i$-th class.

\subsection{Bayes classifier}

An exact solution to this problem is provided by the Bayes classifier. To simplify the discussion we will make the following assumption.

\begin{assumption} \label{ass:1}
 For all $x$ in the feature space we have $d_i(x) > 0$.
\end{assumption}
Commonly used classifiers such as linear discriminant analysis, (penalized) logistic regression or support vector machines all satisfy this assumption.

The Bayes classifier predicts that the probability of $x$ belonging to the $i$-th class is
\begin{align}
 p^\Pi_i(x) = \frac{d_i(x) P_i}{\sum_{k=1}^K d_k(x)P_k}.
\end{align}

Besides the multi-class classifier we may also consider the binary Bayes classifiers for any pair $(i,j)$ of classes. Assuming prior on the two classes is proportional to $(P_i, P_j)$, the binary Bayes classifier predicts the probability of $i$-th class as

\begin{align}
	p_{ij}^\Pi(x) = \frac{d_i(x) P_i}{d_i(x)P_i + d_j(x)P_j}.
\end{align}

Knowledge of vector $\boldsymbol{p}^\Pi(x)$ is  equivalent to knowing the functions $p^\Pi_{ij}(x)$. Indeed under Assumption \ref{ass:1} we have
\begin{align}
p_{ij}^\Pi(x) = \frac{p_i^\Pi(x)}{p_i^\Pi(x) + p_j^\Pi(x)}. \label{eq:bt1}
\end{align}

Conversely, it is easy to show that  the system of equations \eqref{eq:bt1} has a unique solution.

\begin{prop} \label{prop:binary2multi}
The system of  equations \eqref{eq:bt1} in unknowns $p_i^\Pi(x)$ has a unique solution which is the multi-class Bayes classifier.
\end{prop}
%\begin{proof}
%Indeed, we have 
%\begin{align*}
%\frac{1}{r_{ij}(x)} - 1 = \frac{p_j^\Pi}{p_i^\Pi}
%\end{align*}
%\end{proof}

\subsection{One-vs-one classification }

In practice one does not have know true probability distributions and hence cannot construct Bayes classifier. However, given a sample from the distribution, we can train binary classifiers to provide approximations $r_{ij}(x)$ to pairwise Bayes classifiers $p_{ij}^\Pi(x)$. The goal of probabilistic one-vs-one classification is to deduce for any $x$ in the feature space $X$ an estimate $\hat{\boldsymbol{p}}(x)$ for $\boldsymbol{p}^\Pi(x)$. One does it by finding an approximate solution of the system of equations
%\begin{align}
%
%\end{align}
%
%
%Suppose that $K$ classes $C_1, \ldots, C_K$ are distributed according to probability distributions $p_i$ on a  space $X$. Let us denote the $r_{ij}(x)$ the output of Bayes (binary) classifier for $x$ in $X$ yielding the probability $p(x \in C_i \mid x\in C_i \cup C_j )$. Then we have
%\begin{equation}
%r_{ij}(x)= \frac {p_i(x)}{p_i(x) + p_j(x)}.
%\end{equation}
%
%
%The system of equations is called Bradley--Terry equations.  The equations in \eqref{eq:bt1}  can be transformed to a system of linear equations. The theoretical basis for one-vs-one classification framework is provided by the following result.
%
%
%
%Suppose we have estimators $\hat r_{ij}$ for Bayes predictions $r_{ij}$ for all pairs $i\not= j$. A \emph{coupling method} obtains a multi-class probabilistic estimate $\boldsymbol{\hat p}= (\hat p_1, \ldots, \hat p_K)$ that satisfy
\begin{equation}
	{r}_{ij}(x) \approx \frac {\hat p_i(x)}{\hat p_i(x) + \hat p_j(x)}. \label{eq:bt2}
\end{equation}

% One should note that 
%\begin{itemize}
%\item the resulting system of equations will usually not be consistent, because $\hat{r}_{ij}$ are only estimates of true values $r_{ij}$,
%\item when probabilities are parametrized in other ways (e.g. other common parametrizations are as odds, or log-odds), the system of equations \eqref{eq:bt} is non-linear.
%\end{itemize}

Equations forming system \eqref{eq:bt2} are called Bradley-Terry equations.
The system is usually inconsistent and additional assumptions are needed. A method to solve Bradley-Terry equations is called a coupling method. Its input can be represented by a matrix $\boldsymbol{R}(x)$ with off-diagonal entries being $r_{ij}(x)$. Schematically, classification using coupling method $\boldsymbol{v}$ proceeds as indicated in Figure \ref{fig:coupling}.

\begin{figure}[!h]
	\centering
	\begin{tikzcd}[column sep=3cm]
		\textrm{sample $x$}\arrow[r, "binary~classifiers"] \arrow[dr,dotted,swap, "multi-class~ prediction"] & \boldsymbol{R}(x) %\arrow{d}{coupling}[swap]{\boldsymbol{v}} \\
		\arrow{d}{coupling} \\
		 & \boldsymbol{v}(\boldsymbol{R}(x)) 
	\end{tikzcd}
	\caption{Multi-class classification using a coupling method $\boldsymbol{v}$}
	\label{fig:coupling}
\end{figure}

\subsection{Exact desiderata on coupling method} \label{sec:des:exact}

Let us outline what kind of conditions may be desirable in a coupling method. In %Section \ref{sec:des:exact} 
this section
 we will describe exact mathematical criteria, while in section \ref{sec:des:inexact} we shall describe more flexible properties.

%\begin{itemize}
%\item a coupling method should should always provide a unique solution, at least when all $r_{ij}>0$,
%\item Bradley-Terry consistency,
%\item canonicity,
%\item symmetry,
%\item Bayes covariance,
%
%\item good accuracy on benchmark tasks,
%\item lack of parameters, or at least having a small number of parameters,
%\item Hinton's minimality,
%\item simplicity of computation.
%\end{itemize}

\emph{Unique solution} refers to the fact that the procedure to compute a multi-class probabilistic prediction always converges to a unique solution. One way to satisfy this condition would be that it amounts to a linear system of equations in $K$ variables, and the matrix of the system would be invertible. Another example would be when the procedure amounts to finding the minimum of a strongly convex function.

\emph{Bradley-Terry consistency} refers to the outcome of a procedure providing a unique solution. The requirement states that should we apply the coupling procedure to binary Bayes classifiers, then the result is the output of multi-class Bayes classifier i.e. satisfies \eqref{eq:bt1}.

\emph{Canonicity} refers to requirement that the method should be natural. Borrowing an example from regression analysis, Gauss-Markov theorem states that ordinary least squares (OLS) estimates is best linear unbiased estimate of the coefficients of a linear model. Therefore OLS  is a \emph{canonical} way to estimate parameters of a regression model.

Note that potentially there may be multiple canonical methods, derived from (or satisfying) differing sets of assumptions. For instance, in the theory of scoring rules, both logarithmic and quadratic scores can be viewed as canonical [\cite{shannon1948mathematical,selten1998axiomatic}].

\emph{Symmetry} refers to the action of the permutation group which acts both on the classes as well as on multi-class predictions. The action of a permutation $\sigma$ extends to the action on the pairwise matrix $\boldsymbol{R} = (r_{ij})$ by 
\begin{equation}
	 r^\sigma_{i,j} \stackrel{def}{=}  r_{\sigma(i), \sigma(j)}.
\end{equation}
We say that a coupling method $V$ is \emph{symmetric} if 
\begin{align}
		\boldsymbol{v}(\boldsymbol{R})^\sigma = \boldsymbol{v}(\boldsymbol{R}^\sigma)\quad\textrm{for any permutation $\sigma$}.
\end{align}

\emph{Bayes covariance} is a notion introduced in  [\cite{vsuch2016bayes}]. It refers to the behavior of a probabilistic multi-class method built in one-vs-one fashion, when the priors change. Any Bayes classifier, whether binary or multiclass changes its prediction when the priors change. Let us describe the process. Suppose we change priors from $\Pi = (P_1, \ldots, P_k)$ to $\Pi'= (P'_1, \ldots, P'_K)$. Then the prediction of Bayes classifiers changes from $\boldsymbol{p}^\Pi= (p_1, \ldots, p_K)$ to the vector 
\begin{align}
\boldsymbol{p}^{\Pi'} \propto \biggl(p_1 \frac{P'_1}{P_1}, \ldots, p_K \frac{P'_K}{P_K}\biggr). \label{eq:changePrior}
\end{align}

When we use a coupling method and the prior changes, we  can apply \eqref{eq:changePrior} either before or after coupling. If the results are identical, for all initial pairwise data $\boldsymbol{R}$, then we say that the coupling method is \emph{Bayes covariant}.

Let us make this explicit. Let $\boldsymbol{v}$ be a coupling method. Denote by  $\pi$ the change of priors $\pi:\Pi \rightarrow \Pi'$. If we have a  binary classifier predicting distribution $D \propto (p, p')$ on two classes $C_i$ and $C_j$, then we define new distribution $\pi_2^{ij}(D)$ by requiring
$$
\pi_2^{ij}(D) \propto \biggl(p \frac{P'_i}{P_i},p' \frac{P'_j}{P_j}\biggr).
$$
We denote by $\boldsymbol{R}^\pi$ the $K\times K$ matrix with entries $\pi_2^{ij}(r_{ij})$. The value $\boldsymbol{v}(\boldsymbol{R}^\pi)$ represents the effect of change of priors if we apply \eqref{eq:changePrior} \emph{before} coupling.

On the other hand we can apply \eqref{eq:changePrior} after coupling. Say we have a distribution on $K$ classes
\begin{align}
D \propto (p_1, \ldots, p_K)
\end{align}
then we define $D^\pi$ by requiring 
\begin{align}
D^\pi \propto \biggl(p_1 \frac{P'_1}{P_1}, \ldots, p_K \frac{P'_K}{P_K}\biggr).
\end{align}
We now say that a method if Bayes covariant if
\begin{align}
\boldsymbol{v}(\boldsymbol{R}^\pi)=	\bigl(\boldsymbol{v}(\boldsymbol{R})\bigr)^\pi,
\end{align}
for any matrix $\boldsymbol{R}$ with $r_{ij}> 0$.

%Let $\boldsymbol{R}^\pi$ be the matrix of pairwise predictions when we apply \eqref{eq:changePrior} to each binary classifier.
%For Bayes covariant method $\boldsymbol{v}$ the following diagram commutes.
%
%\begin{figure}[!h]
%\centering
%\begin{tikzcd}
%	\boldsymbol{R} \arrow[r, "\pi"] \arrow[d, "coupling"] & \boldsymbol{R}^\pi \arrow[d, "coupling"] \\
%	\boldsymbol{v}(\boldsymbol{R}) \arrow[r, "\pi"] & \boldsymbol{v}(\boldsymbol{R}^\pi) =\boldsymbol{v}(\boldsymbol{R})^\pi
%\end{tikzcd}
%\caption{Diagrammatical description of Bayes covariance for a coupling method $\boldsymbol{v}$}
%\label{fig:bc}
%\end{figure}

\subsection{Inexact desiderata for coupling methods}

\label{sec:des:inexact}

In this section we describe several concepts that lack strict mathematical delineation, but may still be important to consider.

\emph{Good accuracy on benchmark tasks} is  a strong argument to prefer a particular coupling method. This criterion is of course subjective, because there is a great variety of datasets in practice and inevitably there will be instances where any coupling method would underperform. However a good coupling method should perform robustly across a variety of datasets. A historical  example to follow is linear discriminant (LDA) of  Sir Ronald Fisher. At the time he was submitting his paper describing LDA, he had confirmed usefulness of his method not only on the famous iris dataset, but also on two separate collaborations with archaeologists.

\emph{Lack of parameters} is strongly beneficial for situations when the dataset is small and it would be imprudent to reserve a sizeable portion to obtain unbiased estimates of the parameters of the coupling method. This issue gains expediency when the underlying binary classification method requires crossvalidation to select some hyperparameters. For example, this is the case of support vector machines, where often one opts for RBF kernel, which requires choosing two hyper-parameters.

\emph{Simplicity of computation}  may refer to the computational requirement for inference. With the dramatic increase of edge computing devices, such as smartphones, a simple, fast, energy-efficient algorithm may be preferable to a better performing one which would require more computational resources.

Let us finally mention \emph{Hinton's minimality}. Geoffrey Hinton is reported to object to the underlying principle of one-vs-one classification, where any single classifier influences the multi-class prediction [\cite[p.~467]{hastie1998classification}]. Suppose we are classifying objects from CIFAR-10 dataset.  If the true class is  a ``dog'', why should we care what is the output of the classifier distinguishing ``airplanes'' from ``ships''? The binary classifier is trained only on airplanes and ships, and has never seen a dog. One may thus prefer a classifier which somehow suppresses noisy output of (seemingly) irrelevant classifiers.

%The first that the five conditions have a clear mathematical definition, whereas the rest are not exact. For instance while the ideal situation for a coupling method is to have no parameters, the presence of one or two trainable parameters may be justified, if it yields noticeably better performance.

\section{Examples of coupling methods} \label{sec:coupling}

There is a plethora of methods available for coupling. In this section we describe four major classes.

\subsection{Methods aiming to minimize binary divergence}

A multi-class estimate $\hat{\boldsymbol{p}}$ obtained by a coupling method implies via \eqref{eq:bt1} probability distributions on each pair of classes:
\begin{align*}
	\hat D_{ij}^\textrm{multi}= \biggl(\frac{\hat{p_i}}{\hat p_i + \hat p_j},\frac{\hat{p_j}}{\hat p_i + \hat p_j}\biggr)	
\end{align*}
On the other hand, binary classifiers provide another distribution for each pair of  classes, namely
\begin{align*}
\hat D_{ij}^\textrm{binary} = (\hat r_{ij}, \hat r_{ji}).
\end{align*}

A natural way to find  $\hat{\boldsymbol{p}}$ is thus to minimize a functional incorporating divergence measures between these distributions
\begin{align*}
\hat{\boldsymbol{p}} \stackrel{def}{=} \argmin_{\boldsymbol{p}} \sum_{ij} w_{ij} \divm (\hat D_{ij}^\textrm{multi}, \hat D_{ij}^\textrm{binary})
\end{align*}

A notable example which uses this principle is the coupling method introduced in [\cite{hastie1998classification}]. Their method uses Kullback-Leibler divergence $\divm_\textrm{KL}$
\begin{align*}
	\divm\nolimits_\textrm{KL} (\boldsymbol{p}, \boldsymbol{q})= \sum p_i \log (p_i/ q_i)
\end{align*}

Kullback-Leibler divergence strongly penalizes  mispredictions when the true class is predicted to have very low probability. In the context of classification, when we are primarily concerned with accuracy , one could advantageously use other Bregman divergence derived from a proper score [\cite{gneiting2007strictly, buja2005loss}].

For instance, a popular alternative is divergence arising from the quadratic  score [\cite{gneiting2007strictly}] (also known as Brier score [\cite{brier1950verification}]). This divergence takes the form
$$
\divm\nolimits_\textrm{quad} (\boldsymbol{p}, \boldsymbol{q})= \sum (p_i - q_i)^2.
$$
Unlike Kullback-Leibler divergence, the penalty for incorrect prediction with arbitrarily low probability is bounded (by one). This quadratic score is the basis of the popular Wu-Lin-Weng's method [\cite{wu2004probability}].

Their method uses an additional trick, which we term \emph{self-referentiality}. In the method of Hastie and Tibshirani, $w_{ij}$ are constants derived from observed class frequencies. However, we could use weights $w_{ij}$ that increase when the predicted probability of either class $i$ or $j$ is high. A reasonable choice would be for instance $w_{ij}= p_i + p_j$. In the case of quadratic score, it is more convenient to use weights $w_{ij} = (p_i + p_j)^2$ because then the resulting optimization problem is just minimization of a quadratic form
\begin{align*}
\hat{\boldsymbol{p}} \stackrel{def}{=} \argmin_{\boldsymbol{p}} \sum_{i,j} (r_{ij}p_j - r_{ji}p_i)^2.
\end{align*}

We note that the method is symmetric. The choice of quadratic scoring rule can also be viewed as canonical in view of axiomatic characterization given in  [\cite{selten1998axiomatic}].  However the choice of weights $w_{ij}$ to make the method computationally amenable is somewhat arbitrary. Thus we do not view Wu-Lin-Weng's method as canonical.

\subsection{Arboreal methods}

System of equations \eqref{eq:bt2} is overdetermined because there are $\binom{K}{2}$ equations together with the requirement that the total sum of predicted probabilities is one.  Given an overdetermined system of equations, a common approach is to select a minimal subset of equations needed to solve for the unknowns. This principle underlies arboreal coupling methods.

An arboreal coupling method ignores all but $K-1$ of the pairwise predictions. A simple example for 3 classes is the classifier which ignores prediction of the binary classifier distinguishing between classes 1 and 3. The two equations form \eqref{eq:bt2} together with the requirement that the prediction probabilities should sum to 1 yield the following system of equations
\begin{equation}
	\begin{split}
		\frac{\hat p_1}{\hat p_1 + \hat p_2} &= {r}_{12}\\
		\frac{\hat p_2}{\hat p_2 + \hat p_3} &= {r}_{23}\\
		\hat p_1 + \hat p_2 + \hat p_3 &= 1
	\end{split}
	\label{eq:arb1}
\end{equation}

For $K=3$, there are three different arboreal coupling methods, which we shall call $\boldsymbol{s}_1$, $\boldsymbol{s}_2$, $\boldsymbol{s}_3$, where the system \eqref{eq:arb1} represents the coupling method $\boldsymbol{s}_2$.

In order to obtain a regular system of equations for  larger values of $K$ , the resulting classifiers must represent a connected tree within the complete graph on all classes. Any arboreal method is non-canonical, because there is no natural way to choose a tree in a complete graph.

\subsection{Methods emphasizing Hinton's minimality}

A variant on arboreal methods is the following oracle method which we shall call Hinton's oracle. It  assumes it has the access to the ground truth $y = C_t$. It outputs probabilities by solving only the equations
$$
\frac{\hat p_t}{\hat p_t + \hat p_j} = {r}_{tj},\quad \textrm{for }j\not = t,
$$
together the  with total sum equation
$$
\sum_{i=1}^K  \hat p_i = 1.
$$

For any sample, its result is identical to the arboreal method where the underlying tree forms a star. This method is clearly the optimal one with respect to Hinton's minimality criterion, because it ignores the output of any classifier not trained on the true class. It is not a true classification method, however, because it ``peeks'' at the label. Its main utility is to provide a way to benchmark other methods with respect to Hinton's minimality.

In order to convert Hinton's oracle into a true classification method we can use the principle of self-referentiality.  Denote by $\boldsymbol{s}_i$ the prediction of the arboreal method, where the underlying tree is a star centered at class $C_i$. The predictions will generally span the vector space $R^K$. In that case the prediction $\hat{\boldsymbol{p}}$ of the method  can be expressed as a linear combination of $\boldsymbol{s}_i(x)$. Let us now define \emph{radial} coupling method. The  method requires that for the  weights we could in fact use the components of $\boldsymbol{p}$ i.e.
\begin{equation}
	\begin{split}
	\hat {\boldsymbol{p}} &= \hat p_1 \boldsymbol{s}_1(x) + \ldots + \hat p_K \boldsymbol{s}_K(x)\\
	1 &= \hat p_1 + \ldots + \hat p_K
	\end{split}
	 \label{eq:radial}
\end{equation}
%
The intuition behind \eqref{eq:radial} is that if some component of $\hat{\boldsymbol{p}}$ is close to 1 (and thus the remaining components are close to zero), then the prediction $\hat{\boldsymbol{p}}$ should be close to the prediction of Hinton's oracle which is provided by ${{T}}_i$.

\subsection{Bayes covariant methods}

There are many Bayes covariant methods. First examples are provided by arboreal coupling methods.

\begin{prop}
Any arboreal method $\boldsymbol{t}$ is Bayes covariant.
\end{prop}
\begin{proof}
We start by defining property $BC_{ij}$ (short for Bayes covariant when restricted to classes $i$ and $j$) of a coupling method. A coupling method $\boldsymbol{v}$ satisfies $BC_{ij}$ if for any change of priors $\pi$ and any  pairwise data $\boldsymbol{R}$ the odds of classes $C_i$ to $C_j$ predicted by $\boldsymbol{v}(\boldsymbol{R}^\pi)$ and $\bigl(\boldsymbol{v}(\boldsymbol{R})\bigr)^\pi$ are equal.

Let us list some easy to verify properties of this concept:

\begin{itemize}
\item[a)] a Bayes covariant coupling method satisfies $BC_{ij}$ for any $i\not= j$,
\item[b)] if a coupling method satisfies $BC_{12}, BC_{23}, \ldots, BC_{(K-1),K}$ then it is Bayes covariant,
\item[c)] if a coupling method satisfies $BC_{ij}$ for all $i\not= j$ then it is Bayes covariant,
\item[d)] if a coupling method satisfies $BC_{ij}$ and $BC_{jk}$ then it satisfies $BC_{ik}$.
\end{itemize}

An arboreal coupling method $\boldsymbol{t}$ induced by a spanning tree $T$ on the classes satisfies $BC_{ij}$ whenever the edge $C_iC_j$ belongs to $T$. It follows using from b) and d) that the method is Bayes covariant.

%	
%	
%Let $\pi$ be a change of priors from $\Pi = (P_1, \ldots, P_K)$ to $(P'_1, \ldots, P'_K)$. We need to verify $\boldsymbol{v}(\boldsymbol{R}^\pi) =\bigl( \boldsymbol{v}(\boldsymbol{R})\bigr)$ for any pairwise matrix $\boldsymbol{R}$. |We will rephrase this in an equivalent way. Let 
%\begin{align}
%	\begin{split}
%	\boldsymbol{v}(\boldsymbol{R}^\pi) &= (v'_1, \ldots, v'_K) \\
%	\boldsymbol{v}(\boldsymbol{R}) &= (v_1, \ldots, v_K)
%	\end{split}
%\end{align}
%Suppose $\pi$ is a change of priors from $\Pi = (P_1, \ldots, P_K)$ to $\Pi' = (P'_1, \ldots, P'_K)$. If the prediction given $\Pi$ is $(t_1, \ldots, t_K)$ then we need to prove that
%\begin{align*}
%\boldsymbol{t}(R^\pi) = \bigl(t_1 \frac{P'_1}{P_1}, \ldots, t_K \frac{P'_K}{P_K}\bigr)
%\end{align*}
%	
%	We can proceed by induction. The statement is clear when $K=2$. Suppose the proposition is true for an integers $<K$ and we are given a spanning tree $T$ on $K$ vertices. Any tree has a vertex of degree one. We may assume that in the case of $T$ it is the vertex corresponding to $C_1$ and the unique edge incident to $C_1$ is $C_1C_2$.
%	
%	 Denote by $\boldsymbol{t}$ the prediction corresponding to $T$ and by $\boldsymbol{t}'$ the prediction corresponding to the subgraph without the vertex $C_1$. Let us write 
%	\begin{align*}
%	\boldsymbol{t}(x) &= (t_1, t_2, \ldots, t_K) \\
%	\boldsymbol{t'}(x) &= ( t'_2, \ldots, t'_K) 
%\end{align*}
%These vectors are related as follows:
%\begin{align}
%\begin{split}
%\frac{t_1}{t_1+t_2} &= r_{12} \\
%(t_2, \ldots, t_K) &\propto (t'_2, \ldots, t'_{K})
%\end{split} \label{eq:ind1}
%\end{align}
%
%Now consider the effect of change of priors $\Pi = (P_1, \ldots, P_K)$ to $\Pi' = (P'_1, \ldots, P'_K)$. Applying \eqref{eq:changePrior} to the multi-class prediction yields a vector 
%$$
%\boldsymbol{t}^{\Pi\rightarrow \Pi'} \propto (t_1 \frac{P'_1}{P_1}, \ldots, t_K \frac{P'_K}{P_K}).
%$$
%Now consider what happens when we apply \eqref{eq:changePrior} before coupling. By induction assumption the method $\boldsymbol{t}'$ is Bayes covariant and thus applying coupling \emph{after} change of priors yields
%$$
%\boldsymbol{t}'^{\Pi\rightarrow \Pi'} \propto (t_2 \frac{P'_2}{P_2}, \ldots, t_K \frac{P'_K}{P_K})
%$$

\end{proof}

Given several Bayes covariant methods one can form  their ``linear combination'' - an ensemble which is also Bayes covariant. Let us make it explicit.

Suppose real numbers $\alpha_1, \ldots,\alpha_M$ are real numbers and $\boldsymbol{v}_1, \ldots, \boldsymbol{v}_M$ are probabibilistic multi-class classification methods satisfying Assumption \ref{ass:1}. We define their linear combination $\bigoplus_i a_m \boldsymbol{v}_m$ as the classification method that for any $x$ yields the probability distribution on $K$ classes that satisfies
\begin{align*}
 	\bigl(a_1 \boldsymbol{v}_1 \oplus \ldots \oplus a_M \boldsymbol{v}_M\bigr)(x) \propto \boldsymbol{v}_1^{a_1}(x) \odot \cdots \odot \boldsymbol{v}_M^{a_M}(x),
\end{align*}
where the symbol $\odot$ denotes componentwise multiplication, and exponentials are computed component-wise. With this definition we can state the following result.

\begin{prop}
Suppose the coupling methods $\boldsymbol{v}_1, \ldots, \boldsymbol{v}_M$ are Bayes covariant. If $\sum a_m = 1$, then $\bigoplus_m a_m \boldsymbol{v}_m$ is also a Bayes covariant coupling method.
\end{prop}

\begin{proof}
We will be using property $BC_{ij}$ from the proof of the previous proposition. It is sufficient to show that $BC_{ij}$ holds for any $i\not= j$.

Let $\pi$ be a change of priors from $(P_1, \ldots, P_K)$  to $(P'_1, \ldots, P'_K)$. Let $o_m$ denotes the odds of classes $C_i$ and $C_j$ predicted by coupling method $\boldsymbol{v}_m$. Then the odds of $C_i$ and $C_j$ predicted by $(\bigoplus_i a_i \boldsymbol{v}_i)^\pi$ are 
\begin{align}
	\biggl(\bigoplus_m a_m \boldsymbol{v}_m\biggr)^\pi(\boldsymbol{R}) & = \frac{P'_i}{P_i} \frac{P_j}{P'_j} \biggl(\bigoplus_m a_m \boldsymbol{v}_m\biggr)(\boldsymbol{R}) \\
	&= \frac{P'_i}{P_i} \frac{P_j}{P'_j} \prod_m o_m^{a_m}
\end{align}

Let us now use Bayes covariance of $\boldsymbol{v}_i$. We have
\begin{align}
	\biggl(\bigoplus_m a_m \boldsymbol{v}_m\biggr)(\boldsymbol{R^\pi}) & =  \biggl(\bigoplus_m a_m \boldsymbol{v}_m^\pi\biggr)(\boldsymbol{R}) \\
	&= \frac{P'_i}{P_i} \frac{P_j}{P'_j} \prod_m o_m^{a_m}
\end{align}

It follows that the odds of $C_i$ to $C_j$ for $\bigoplus_m a_m \boldsymbol{v}_m$ are 
\begin{align}
\biggl(\frac{P'_i}{P_i} \frac{P_j}{P'_j} \biggr)^{\sum a_m} \prod_m o_m^{a_m}.
\end{align}
Clearly, under the assumption $\sum_m a_m = 1$, the classifier $\bigoplus_m a_m \boldsymbol{v}_m$ satisfies $BC_{ij}$ and it follows that it is Bayes covariant.
\end{proof}

% 
% given such $\alpha + \beta = 1$. If the prediction of the two Bayes covariant methods $\boldsymbol{B}$ and $\boldsymbol{\tilde B}$ are
%\begin{align*}
%	\boldsymbol{B}(x) = (p_1, \ldots, p_K)\quad\textrm{and}\quad \boldsymbol{\tilde B}(x) = (\tilde p_1, \ldots, \tilde p_K)
%\end{align*}
%then the prediction of the ensemble $\alpha 	\boldsymbol{B} \oplus \beta \boldsymbol{\tilde B}$ is given by
%\begin{align*}
%	\bigl(\alpha 	\boldsymbol{B} \oplus \beta \boldsymbol{\tilde B}\bigr)(x)\propto (\,p_1^\alpha \tilde p_1^\beta, \ldots, p_K^\alpha \tilde p_K^\beta).
%\end{align*}

As an example, consider the classification problem with three classes. We have three arboreal classifiers $\boldsymbol{s}_1, \boldsymbol{s}_2, \boldsymbol{s}_3$ which correspond to star graphs centered respectively at classes $C_1, C_2, C_3$. The classifier 
\begin{align}
\frac13 \boldsymbol{s}_1 \oplus \frac13 \boldsymbol{s}_2 \oplus \frac 13 \boldsymbol{s}_3  \label{eq:bc1}
\end{align}
is Bayes covariant \emph{and} symmetric. A key result proved in [\cite{vsuch2016bayes}] is that it is a unique such classifier for $K=3$.

\begin{thm} \label{thm:K3}
	There exists a unique symmetric Bayes covariant coupling methods for $K=3$.
\end{thm}

Let us point out one consequence of the theorem. We could define another ensemble $\boldsymbol{E}$ by averaging the predictions of $\boldsymbol{s}_1, \boldsymbol{T}_2, \boldsymbol{T}_3$:
\[
\boldsymbol{E} = \frac13 \boldsymbol{s}_1(x) + \frac13 \boldsymbol{s}_2(x) + \frac 13 \boldsymbol{s}_3(x).
\]
Note that here we use the plus symbol to denote the vector addition of vectors representing $K$-class probability distributions. This is a different classifier from the one given by  \eqref{eq:bc1} (see Appendix \ref{app:explicit}), and since it is symmetric, it cannot be Bayes covariant.

\subsection{Properties of coupling methods}

The following table indicates which properties are enjoyed by each coupling method we described in this section.

\begin{table}[!ht]
\begin{tabular}{cm{2.5cm}m{1.5cm}m{1.5cm}ccm{1.5cm}m{1.5cm}}
&property & Hastie-Tibshirani & Wu-Lin-Weng & radial & normal & arboreal & Hinton's oracle \\
\hline 
\multirow{5}{*}{\begin{turn}{90}\makecell{exact}\end{turn}}
&uniqueness &  yes & yes & yes & yes & yes & yes \\
&Bradley-Terry consistency & yes & yes & yes & yes & yes & yes \\
&canonicity & yes & no & yes & yes & no & yes \\
&symmetry & yes & yes & yes & yes & no & yes \\
& Bayes covariant & no & no & no & yes & yes & yes \\
\hline
\multirow{5}{*}{\begin{turn}{90}\makecell{non-exact}\end{turn}}
&Hinton's minimality & ? & ?  & ++ & ?  & ? & +++ \\
&lack of parameters & yes & yes & yes & yes & yes & yes \\
& amounts to a linear system & no & yes & yes & yes & yes & yes\\
& linear system of special form & no & no & yes & yes & yes & yes \\
& iterative method & yes & yes & yes & no & not needed & not needed\\
\hline
\end{tabular}
\caption{Summary of properties of coupling methods described in Section \ref{sec:coupling}.}
\label{tab:summaryCoupling}
\end{table}

\section{Experimental results.}

The choice of a coupling method is an important task for practitioners. We aim to help by posing the following interrelated questions

\begin{itemize}
\item[a)] Which non-parametric coupling methods shows good performance across diverse data sets? Is it the orthodox method of Wu-Lin-Weng?
\item[b)] If the answer to a) is a coupling method which is not Bayes covariant, can we find a parametric method Bayes covariant method which  matches its performance?
\item[c)] If the answer to b) is negative, is it better to apply change of priors before or after coupling? Additionally, how serious is the impact of the lack of Bayes covariance in practice?
\end{itemize}

These questions will be put to test in Sections \ref{sec:exp1}, \ref{sec:exp2}, and \ref{sec:exp3}.

\subsection{Datasets}

It is natural to consider the benchmark datasets used in previous evaluation of coupling methods [\cite{wu2004probability}]. The authors considered altogether seven datasets with number of classes ranging from 3 to 26. For each dataset they extracted 40 pairs of training and testing datasets. Half of these were smaller (300 training samples and 500 testing samples) and the other half was larger (800 training samples and 1000 testing samples). In our experiments we use only the larger pairs.

As a preliminary step we conducted analysis of linear separability of the datasets. We used ECOS solver in CVXR library to solve the underlying quadratic program. The results are reported in Table \ref{tab:sep}.

\input{tab-sep.tex}

We can infer that except for 'waveform' dataset, majority of binary problems are linearly separable.

\subsection{Classification methodology}

We have opted to use LDA as a binary classification tool. Its appeal is that it is applicable even for linearly separable datasets, thus obviating additional crossvalidation step incurred by methods that use regularization (e.g. penalized logistic regression or SVM). It usually provides results close to logistic regression [\cite{james2013introduction}].

For preprocessing we first removed features that were constant across classes. Then we projected the data using PCA to the subspace for which corresponding singular values were at least $1/1000$. We used PCA without scaling or centering since the datasets' features were already normalized to interval $[-1,1]$.

For training LDA we used MASS library in R. We implemented coupling methods in C++ using Eigen library. As arguments to the coupling methods we use log-odds, i.e. instead of $r_{ij}$ they expect
$$
\tilde r_{ij} = \log \biggl(\frac{1}{r_{ji}} - 1\biggr).
$$
The resulting matrix $\tilde{\boldsymbol{R}}$ is skew symmetric (with the convention of having zeroes on the diagonal).

This parametrization allows for precise representation of extremely small probabilities which are sometimes produced by LDA. These small probabilities might be expected given that a large fraction of binary problems is linearly separable.

Small probabilities generally have very little effect on the implementation of Bayes covariant methods which generally operate using log-odds. However small probabilities do affect both the method of Wu-Lin-Weng as well sa radial method, because then Assumption \ref{ass:1} fails to hold.

%\begin{figure}[!ht]
%	\includegraphics[width = 0.5\linewidth]{graph/dna-lda.pdf}
%	\includegraphics[width = 0.5\linewidth]{graph/waveform-lda.pdf}
%	\caption{Projections of three-class datasets using LDA Left: dna, right: waveform}
%\end{figure}

\subsection{Experiment 1: Accuracy of parameter-less coupling methods on benchmark datasets} \label{sec:exp1}

The first experiment compares the performanc of three parameterless coupling methods
\begin{itemize}
\item method of Wu-Lin-Weng, which is widely used in software libraries, and thus can be viewed as an orthodox choice,
\item normal coupling, which is a Bayes covariant method,
\item radial coupling, which is not Bayes covariant.
\end{itemize}

Although previous evaluations of the latter two methods were done in [\cite{vsuch2016bayes}], they used phonetic data for which ground truth may be imprecise.

\input{tab-multi.tex}

From Table \ref{tab:multi} we can clearly see that Wu-Lin-Weng's method has the best performance. On some datasets it trounces normal coupling method by a wide margin. The edge over radial method is smaller, with the exception of usps dataset where its accuracy is higher by remarkable 30\%.

\subsection{Experiment 2: Accuracy of Bayes covariant methods with parameters}
 \label{sec:exp2}

In the previous experiment we established that the method of Wu-Lin-Weng provides leading performance. However in view of Theorem \ref{thm:K3} the method is not Bayes covariant.

In this experiment we investigated whether we can (always) find  a Bayes covariant coupling method that could match the performance of Wu-Lin-Weng's method, if we allow it to have (a few) trainable parameters.  Our experiment proceeds in five steps.

First, we restrict our investigation to the case of three class datasets. This is a natural starting point, since it is the smallest number of classes where the question is meaningful to ask.

>>>>

\newtheorem{prop}{Proposition}
\newtheorem{thm}{Theorem}
\newtheorem{lem}{Lemma}
\newtheorem{assumption}{Assumption}

@@

\input theory.tex

\input coupling.tex

\input experiment.tex
